Using Occupancy Theory and Hierarchical Models to Estimate Otter Population in Southeastern Minnesota

### 2010-02-24 Paper.tex
Added WinBUGS/R section.

\documentclass[12pt]{article}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{graphicx}
\usepackage{multirow}
\usepackage[dvipsnames, table]{xcolor}
\usepackage[all]{xy}

@@

    \subsection{WinBUGS/R}
    All of our statistical analysis was performed in R.

    All of our modeling was performed using BRugs, which allowed us to use
    WinBUGS commands and other stuff in R.

>>>>

\documentclass[12pt]{article}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{graphicx}
\usepackage{multirow}
\usepackage[dvipsnames, table]{xcolor}
\usepackage[all]{xy}
\usepackage{hyperref}

@@

    \subsection{WinBUGS/R}

    To implement a Bayesian hierarchical model, we make use of the software
    packages WinBUGS \cite{Lunn2000}, R \cite{R2009}, BRugs \cite{Thomas2008},
    and, by extension, OpenBUGS \cite{Thomas2006}, as one can describe the model
    using an ``intuitive `pseudo-code' representation'' in WinBUGS
    \cite{MacKenzie2006} and then run the model within the familiar environment
    of R using the combination of BRugs and OpenBUGS. OpenBUGS, the open source,
    cross-platform successor to WinBUGS, uses the same syntax. All of the
    aforementioned software packages are open source and available
    freely\footnote{WinBUGS: \url{http://www.mrc-bsu.cam.ac.uk/bugs/} and R:
    \url{http://www.r-project.org/} are available as standalone programs. The
    BRugs/OpenBUGS package can be installed by typing
    \texttt{install.packages("BRugs")} in R's interactive shell. Note that while
    R and OpenBUGS include official support for Mac OS X and Unix/Unix-like
    operating systems, WinBUGS and the BRugs package are only available for
    Windows.}.

    WinBUGS requires a representation of a model, data, and initial values on
    which a predetermined number of iterations of the MCMC algorithm are
    executed \cite{MacKenzie2006}. Because WinBUGS explicitly estimates missing
    data that is `padded out' by inserting NA values \cite{Spiegelhalter2003},
    we implement a nested indexing strategy to prevent this extraneous
    estimation on the actual data. This is a particularly important
    implementation detail in this case because the data we received from the
    Minnesota Department of Natural Resources was not consistent with respect to
    our covariates. Neither the number of observers nor number of observations
    after each snow event exhibited consistency. Our simulations, on the other
    hand, involve balanced data sets that are much less difficult to implement
    within WinBUGS and R.

    % Should we say anything about making maps/other CAR-related stuff?
